---
title: "hw02-dominic-tu.rmd"
author: "Dominic Tu"
date: "9/28/2017"
output: github_document
---
Problem 1 
---------
```{r}
#inside hw02 data file 
```
Problem 2
--------
```{r}
library(dplyr)
dat = read.csv('nba2017-player-statistics.csv')
dat
str(dat)
dat = mutate(dat, 
             Player = as.character(Player),
             Team = as.character(Team),
             Experience = as.character(Experience))

library(readr)
dat2 = read_csv('nba2017-player-statistics.csv')
dat2
str(dat2)
dat2 = mutate(dat2, Position = as.factor(Position))

```
Problem 3
-------
```{r}
#library(plyr)
dat$Experience = replace(dat$Experience, dat$Experience ==  "R", 0)
dat$Experience = as.integer(dat$Experience)
dat
```
Problem 4
--------
```{r}
dat = mutate(dat, 
             Missed_FG = dat$FGA - dat$FGM, 
             Missed_FT = dat$FTA - dat$FTM,
             PTS = dat$FTM + 2*dat$Points2 + 3*dat$Points3,
             REB = dat$OREB + dat$DREB, 
             MPG = dat$MIN / dat$GP)
dat = mutate(dat, 
             EFF = (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO) / dat$GP)
dat
length(dat$EFF)
summary(dat$EFF)

arrange(select(dat, Player, Team, Salary, EFF),
        desc(EFF))
filter(dat, EFF < 0)

PTS = cor(dat$EFF, dat$PTS)
REB = cor(dat$EFF, dat$REB)
AST = cor(dat$EFF, dat$AST)
STL = cor(dat$EFF, dat$STL)
BLK = cor(dat$EFF, dat$BLK)
Missed_FG = cor(dat$EFF, dat$Missed_FG)
Missed_FT = cor(dat$EFF, dat$Missed_FT)
TO = cor(dat$EFF, dat$TO)
GP = cor(dat$EFF, dat$GP)

correlation = c(PTS = cor(dat$EFF, dat$PTS), 
                REB = cor(dat$EFF, dat$REB), 
                AST = cor(dat$EFF, dat$AST),
                STL = cor(dat$EFF, dat$STL),
                BLK = cor(dat$EFF, dat$BLK),
                Missed_FG = -cor(dat$EFF, dat$Missed_FG),
                Missed_FT = -cor(dat$EFF, dat$Missed_FT),
                TO = -cor(dat$EFF, dat$TO),
                GP = cor(dat$EFF, dat$GP))
sorted_cor = sort(correlation, decreasing = TRUE)
barplot(sorted_cor, main = 'Correlations between Player Stats and EFF')
```
Problem 5
--------
```{r}
plot(dat$EFF, dat$Salary, cex = .8, xlab = "Efficiency", ylab = "Salary (in millions)")
lines(lowess(dat$EFF, dat$Salary), col = 'red')
cor(dat$EFF, dat$Salary)
#There is a positive correlation between players' efficiency and their respective salary. A higher players' efficiency leads to a higher salary.

player2 = dat[dat$MPG >= 20, ]
player2
plot(player2$EFF, player2$Salary, ylab = "Salary", xlab = "Efficiency")
lines(lowess(dat$EFF, dat$Salary), col = 'red')
cor(player2$EFF, player2$Salary)

#There is less of a correlation between efficiency and salary in "more established players". Once a player has more experience, their salary has a lessen correlation to efficiency.  
````
Problem 6
--------
#Comments and Reflection 

> This project was more straight forward than hw01. We saw a lot of the functions and task from labs. The hardest part of the assignment was making sure objects were found. The HW took me about 3 hours. I most time consuming part was problem 4. The most frustating part was knitting the file to find that an object is not found. Overall, I enjoyed learning more about the NBA and breaking it down through a statistical approach. 



