---
title: "hw03-dominic-tu.rmd"
author: "Dominic Tu"
date: "10/11/2017"
output: github_document
---
#Ranking of teams
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
rmarkdown::render("hw03-dominic-tu.rmd")
````
##Basic Rankings
```{r}
teams
ggplot(teams,
       aes(x = reorder(team, salary),
           y = salary)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  ggtitle("NBA Teams ranked by Total Salary") +
  labs(x = "Team", y = "Salary (in millions)") +
  geom_hline(data = teams, aes(yintercept = mean(salary)), color = 'red', alpha = .4, size  = 1)

ggplot(teams,
       aes(x = reorder(team, points),
           y = points)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  ggtitle("NBA Teams ranked by Total Points") +
  labs(x = "Team", y = "Total Points") +
  geom_hline(data = teams, aes(yintercept = mean(points)), color = 'red', alpha = .4, size = 1)

ggplot(teams,
       aes(x = reorder(team, efficiency),
           y = efficiency)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  ggtitle("NBA Teams ranked by Total Efficiency") +
  labs(x = "Team", y = "Total Efficiency") +
  geom_hline(data = teams, aes(yintercept = mean(efficiency)), color = 'red', alpha = .4, size = 1)
```
##PCA
```{r}
pca = prcomp(select(teams, points3, points2, free_throws,
                    off_rebounds,def_rebounds,  assists, steals,
                    blocks, turnovers, fouls), scale. = TRUE)
eigenvalue = pca$sdev^2
df_eigs = data.frame(eigenvalue, 
                     prop = round(eigenvalue/sum(eigenvalue), 4),
                     cumprop = round(cumsum(eigenvalue/10), 4))

ggplot(teams, aes(x = pca$x[ , 1], y = pca$x[ , 2], label = team)) +
  geom_text(size = 3.5) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) + 
  ggtitle("PCA plot (PC1 and PC2)") + 
  labs(x = "PC1", y = "PC2")


s1 = 100*((pca$x[ ,1] - min(pca$x[ ,1])) / (max(pca$x[ ,1]) - min(pca$x[ ,1])))
teams = mutate(teams, PC1 = s1)
teams 

ggplot(teams, aes(x = reorder(team, s1), y = s1)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(x = "Team", y = "First PC (scaled from 0 to 100") +
  ggtitle("NBA Teams ranked by scaled PC1")
```
- The PCA1 ranking shows that the Golden State Warriors show the more variation with (PCA1 = 100), and Sacramento Kings show the least amount of variation with (PCA1 = 0)

-------
#comments and reflection
> I think the hardest part was learning PCA on our own. It was brand new to me. Relative path was quite easy once you learn how to use it. This assignment took me about 5 hours





